Enforced Invariant Confluence

### 2018-04-22 doc/paper/sections/conclusion.tex
Finished paper tweaks.

\section{Conclusion}
TODO

>>>>

\section{Conclusion}
This paper revolved around two major contributions.  First, we developed a
deeper understanding of \invariantclosure{} and \invariantconfluence{} by
looking at the two criteria with reachability in mind. We found that
\invariantclosure{} fails to incorporate a notion of reachability, and using
this intuition, we developed conditions under which \invariantclosure{} and
\invariantconfluence{} are equivalent.  Second, we proposed a new consistency
model and generalization of \invariantconfluence{}, segmented
\invariantconfluence{}, that can be used to replicate non-\invariantconfluent{}
objects with a small amount of coordination while still preserving their
invariants. We also found that segmented \invariantconfluence{} naturally
subsumes existing techniques for maintaining invariants of replicated objects
(e.g.\ locking and escrow transactions).

### 2018-04-22 doc/paper/sections/evaluation.tex
Finished paper tweaks.

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe and evaluate Lucy: a prototype implementation of
our decision procedures and system models.
% We evaluate Lucy by answering the following questions:
% \begin{itemize}
%   \item
%     How practical is the interactive \invariantconfluence{} decision procedure?
%     Can we use it to classify real-world transactions and invariants?
%   \item
%     How practical is segmented \invariantconfluence{}? Are real-world workloads
%     amenable to segmentation?
%   \item
%     How efficient is the interactive \invariantconfluence{} decision procedure?
%   \item
%     How efficiently can we replicate a segmented \invariantconfluent{} object as
%     compared to alternative approaches like replicating with weak or strong
%     consistency?
%   \item
%     How does the performance of replicating a segmented \invariantconfluence{}
%     object vary as we vary the workload, segmentation, and replication factor?
% \end{itemize}

@@

Lucy also includes an implementation of the \invariantconfluence{} and
segmented-invariant confluence system models in roughly 3,500 lines of C++.
Users specify objects, transactions, invariants, and segmentations in C++. Lucy
then replicates the objects using segmented \invariantconfluence{} (or
\invariantconfluence{} if the segmentation contains a single segment without any
disallowed transactions). Clients send every transaction request to a randomly
selected server. When a server receives a transaction request, it executes
\algoref{TxnExecution} to attempt to execute the transaction locally. If the
transaction requires global coordination, then the server forwards the
transaction request to a predetermined leader. When the leader receives a
transaction request, it sends a coordination request to all other servers. When
a server receives a coordination request from the leader, it stops processing
transactions and sends the leader its state in a coordination reply. When the
leader receives coordination replies from all other servers, it executes the
transaction, and then sends its state to the other servers. When a server
receives a new state, it adopts the state, computes its new active segment, and
resumes normal processing. After every 100 transactions processed, a server
sends a merge request to a randomly selected server.
% Merge requests are tagged with a monotonically increasing epoch number that is
% incremented by the master after every round of global coordination. This allows
% servers to discard merge requests from previous epochs.

Lucy can also replicate an object with eventual consistency and with
linearizability. With eventual consistency, clients send every transaction
request to a randomly selected server. The server executes the transaction
locally and returns immediately to the client, sending merge requests after
every 100 transactions. With linearizability, clients send every transaction
request to a predetermined leader. The leader relays the transaction request to
all other servers, and when the leader receives replies from them, it executes
the transaction and replies to the client. This communication pattern mimics
the ``normal operation'' of state machine replication protocols
\cite{lamport1998part, liskov2012viewstamped}.
% In \secref{SegmentedInvariantConfluenceEval}, we compare the performance of
% replicating with segmented \invariantconfluence{} against the performance of
% replicating with eventual consistency and linearizability.

Because fault-tolerance is largely an orthogonal concern to
\invariantconfluence{}, Lucy is implemented without fault-tolerance. It would be
straightforward to add fault-tolerance to Lucy, but it would not affect our
discussions or evaluation, so we leave it for future work.
% Moreover, users currently have to specify their workloads in Python (for the
% decision procedures) and C++ (for the runtime). In the future, we plan on
% removing this redundancy.

@@

\subsection{Segmented Invariant Confluence}%
\seclabel{SegmentedInvariantConfluenceEval}
Now, we evaluate the performance of replicating an object with segmented
\invariantconfluence{} as compared to replicating it with eventual consistency or
linearizability.
% We begin with two benchmarks that demonstrate the same concept: the performance
% of segmented \invariantconfluent{} replication varies with the amount of global
% coordination induced by either (a) performing a transaction that is disallowed
% within a segment or (b) transitioning between segments.

@@

\begin{benchmark}\benchlabel{VaryWithdraws}
Consider again the PN-Counter from \exampleref{CounreachableExample} and the
corresponding transactions, invariants, and single-segment segmentation that
forbids concurrent decrements. We replicate this object on 32 servers
deployed on 32 m5.xlarge EC2 instances within the same availability zone.  Each
server has three colocated clients that issue deposit and withdrawal
transactions. We replicate the object with eventual consistency, segmented
\invariantconfluence{}, and linearizability and measure the system's total
throughput as we vary the fraction of client requests that are withdrawals. The
results are shown in the top of \figref{ThroughputVsGlobalSyncs}.

Both eventually consistent replication and linearizable replication are
unaffected by the workload, achieving roughly 700,000 and 7,000 transactions
per second respectively.
%
% Expectedly, eventually consistent replication significantly outperforms
% linearizable replication because (a) transactions can be sent to any server
% (not just the leader) and (b) servers do not coordinate with each other at all.
%
Segmented \invariantconfluent{} replication performs well for low-withdrawal
workloads and performs increasingly poorly as we increase the fraction of
withdrawal transactions, eventually performing worse than linearizable
replication. For example, with 5\% withdrawal transactions, segmented
\invariantconfluent{} replication performs an order of magnitude better than
linearizable replication; with 50\% withdrawals, it performs as well; and with
100\% withdrawals, it performs two times worse.

These results are expected. Deposit transactions can execute without any
coordination while withdrawal transactions require global coordination. As we
increase the fraction of withdrawals, we increase the amount of coordination
that the system has to perform which in turn drastically decreases the
throughput. These results also offer two insights:
%
First, for low-withdrawal workloads, segmented \invariantconfluent{} replication
achieves a compromise between strong and weak consistency. It guarantees that
invariants are maintained (which is impossible with eventual consistency if the
object is not \invariantconfluent{}) with performance many times better than
strongly consistent replication.
%
Second, segmented \invariantconfluent{} replication is poorly suited to workloads
that require a large amount of coordination. For workloads without much inherit
concurrency (e.g.\ withdraw-mostly workloads), maintaining invariants is best
done with strong consistency. It provides stronger guarantees with better
performance.
\end{benchmark}

@@

\begin{benchmark}
  In this benchmark, we measure the scale-out of segmented \invariantconfluent{}
  replication. We repeat \benchref{VaryWithdraws} with a 10\% withdrawal rate,
  but this time we vary the number of servers we use to replicate our object.
  When we replicate with $n$ servers, we use $3n$ clients (the $3$ colocated
  clients on each server) as part of the workload. The results are shown in
  \figref{VaryNodes}.

  Eventually consistent replication scales perfectly with the number of nodes,
  confirming the results in~\cite{bailis2014coordination}.
  %
  % With eventually consistent replication, servers do not coordinate at all, so
  % they are completely unaffected by the number of servers.
  %
  Linearizable replication, on the other hand, scales up to about 3-5 servers
  before performance begins to decrease. These numbers are consistent with
  typical deployments of state-machine replication protocols like
  Paxos~\cite{chandra2007paxos}.
  %
  % Because all messages are sent to the leader, the leader becomes the
  % bottleneck as the number of servers and clients increases. Moreover, the
  % leader must wait for responses from more servers, increasing the latency of
  % the slowest response which in turn decreases throughput.
  %
  Segmented \invariantconfluent{} replication scales up to about 6-8 servers
  before succumbing to the same scalability bottlenecks as linearizable
  replication.

  These results highlight the importance of coordination avoidance in
  distributed databases. While segmented \invariantconfluent{} replication scales
  out slightly better than linearizable replication, both scale significantly
  worse than eventually consistent replication even for a very low (i.e.\ 10\%)
  withdrawal workload. This demonstrates that even a small amount of
  coordination can significantly reduce the scalability of a system.
\end{benchmark}

>>>>

\section{Evaluation}\seclabel{Evaluation}
In this section, we describe and evaluate Lucy: a prototype implementation of
our decision procedures and system models.

@@

Lucy also includes an implementation of the \invariantconfluence{} and
segmented-invariant confluence system models in roughly 3,500 lines of C++.
Users specify objects, transactions, invariants, and segmentations in C++. Lucy
then replicates the objects using segmented \invariantconfluence{} (or
\invariantconfluence{} if the segmentation contains a single segment without
any disallowed transactions). Clients send every transaction request to a
randomly selected server. When a server receives a transaction request, it
executes \algoref{TxnExecution} to attempt to execute the transaction locally.
If the transaction requires global coordination, then the server forwards the
transaction request to a predetermined leader. When the leader receives a
transaction request, it sends a coordination request to all other servers. When
a server receives a coordination request from the leader, it stops processing
transactions and sends the leader its state in a coordination reply. When the
leader receives coordination replies from all other servers, it executes the
transaction, and then sends its state to the other servers. When a server
receives a new state, it adopts the state, computes its new active segment, and
resumes normal processing. After every 100 transactions processed, a server
sends a merge request to a randomly selected server.

Lucy can also replicate an object with eventual consistency and with
linearizability. With eventual consistency, clients send every transaction
request to a randomly selected server. The server executes the transaction
locally and returns immediately to the client, sending merge requests after
every 100 transactions. With linearizability, clients send every transaction
request to a predetermined leader. The leader relays the transaction request to
all other servers, and when the leader receives replies from them, it executes
the transaction and replies to the client. This communication pattern mimics
the ``normal operation'' of state machine replication protocols
\cite{lamport1998part, liskov2012viewstamped}.

Because fault-tolerance is largely an orthogonal concern to
\invariantconfluence{}, Lucy is implemented without fault-tolerance. It would
be straightforward to add fault-tolerance to Lucy, but it would not affect our
discussions or evaluation, so we leave it for future work.

@@

\subsection{Segmented Invariant Confluence}%
\seclabel{SegmentedInvariantConfluenceEval}
Now, we evaluate the performance of replicating an object with segmented
\invariantconfluence{} as compared to replicating it with eventual consistency
or linearizability.

@@

\begin{benchmark}\benchlabel{VaryWithdraws}
  Consider again the PN-Counter from \exampleref{CounreachableExample} and the
  corresponding transactions, invariants, and single-segment segmentation that
  forbids concurrent decrements. We replicate this object on 32 servers
  deployed on 32 m5.xlarge EC2 instances within the same availability zone.
  Each server has three colocated clients that issue increment and decrement
  transactions. We replicate the object with eventual consistency, segmented
  \invariantconfluence{}, and linearizability and measure the system's total
  throughput as we vary the fraction of client requests that are decrements.
  The results are shown in the top of \figref{ThroughputVsGlobalSyncs}.

  Both eventually consistent replication and linearizable replication are
  unaffected by the workload, achieving roughly 700,000 and 7,000 transactions
  per second respectively. Segmented \invariantconfluent{} replication performs
  well for low-decrement workloads and performs increasingly poorly as we
  increase the fraction of decrement transactions, eventually performing worse
  than linearizable replication. For example, with 5\% decrement transactions,
  segmented \invariantconfluent{} replication performs an order of magnitude
  better than linearizable replication; with 50\% decrements, it performs as
  well; and with 100\% decrements, it performs two times worse.

  These results are expected. Increment transactions can execute without any
  coordination while decrement transactions require global coordination. As we
  increase the fraction of decrements, we increase the amount of coordination
  that the system has to perform which in turn drastically decreases the
  throughput. These results also offer two insights.
  %
  First, for low-decrement workloads, segmented \invariantconfluent{}
  replication achieves a compromise between strong and weak consistency. It
  guarantees that invariants are maintained (which is impossible with eventual
  consistency if the object is not \invariantconfluent{}) with performance many
  times better than strongly consistent replication.
  %
  Second, segmented \invariantconfluent{} replication is poorly suited to
  workloads that require a large amount of coordination. For workloads without
  much inherit concurrency (e.g.\ decrement-mostly workloads), maintaining
  invariants is best done with strong consistency. It provides stronger
  guarantees with better performance.
\end{benchmark}

@@

\begin{benchmark}
  In this benchmark, we measure the scale-out of segmented
  \invariantconfluent{} replication. We repeat \benchref{VaryWithdraws} with a
  10\% decrement rate, but this time we vary the number of servers we use to
  replicate our object.  When we replicate with $n$ servers, we use $3n$
  clients (the $3$ colocated clients on each server) as part of the workload.
  The results are shown in \figref{VaryNodes}.

  Eventually consistent replication scales perfectly with the number of nodes,
  confirming the results in~\cite{bailis2014coordination}.  Linearizable
  replication, on the other hand, scales up to about 3--5 servers before
  performance begins to decrease. These numbers are consistent with typical
  deployments of state-machine replication protocols like
  Paxos~\cite{chandra2007paxos}. Segmented \invariantconfluent{} replication
  scales up to about 6--8 servers before succumbing to the same scalability
  bottlenecks as linearizable replication.

  These results highlight the importance of coordination avoidance in
  distributed databases. While segmented \invariantconfluent{} replication
  scales out slightly better than linearizable replication, both scale
  significantly worse than eventually consistent replication even for a very
  low (i.e.\ 10\%) decrement workload. This demonstrates that even a small
  amount of coordination can significantly reduce the scalability of a system.
\end{benchmark}

### 2018-04-22 doc/paper/sections/related_work.tex
Finished paper tweaks.

\newcommand{\sieve}{\textsc{SIEVE}}
\textbf{RedBlue Consistency and \sieve.}
RedBlue consistency is a consistency model that sits between causal consistency
and linearizability~\cite{li2012making}.
%
With RedBlue consistency, every operation is manually labelled as either red or
blue. All operations are executed with causal consistency, but with the added
restrictions that red operations are executed in a single total order embedded
within the causal ordering.
%
Invariant safety is a sufficient condition for RedBlue consistent objects to be
\invariantconfluent{}. Invariant safety is an analog of \invariantclosure{} for an
operation-based system model (as opposed-to the state based system model that
we use). In~\cite{li2014automating}, Li et al.\ develop techniques to identify
invariant safe operations. Like \invariantclosure{}, invariant safety is a
sufficient but not necessary condition for \invariantconfluence{}. Li et al.\
develop sophisticated techniques for deciding invariant safety that involve
calculating weakest preconditions. These techniques are complementary to our
work and can be used to improve the \invariantclosure{} subroutine used by our
decision procedures.

\textbf{The Homeostasis Protocol.}
The homeostasis protocol~\cite{roy2015homeostasis} uses program analysis to
avoid unnecessary coordination between servers in a \emph{sharded} database
(whereas \invariantconfluence{} targets \emph{replicated} databases). The
protocol generates global invariants, called global treaties, and servers can
operate without coordination so long as they preserve the treaty. When the
treaty is violated, the nodes perform a round of coordination and establish a
new treaty. The protocol guarantees that transactions are executed with
observational equivalence with respect to some serial execution of the
transactions. This means that intermediate states may be inconsistent, but
externally observable side effects and the final database state are consistent.
The observational equivalence guaranteed by the homeostasis is stronger than
the guarantees of \invariantconfluence{}. As a result, there are invariants and
workloads that the homeostasis protocol would execute with more coordination
than a segmented \invariantconfluent{} execution. Still, the homeostasis
protocol's mechanism of establishing invariants and operating without
coordination so long as the invariants are maintained is very similar to our
design of segmented \invariantconfluence{}.

@@

\textbf{Token Based Invariant Confluence.}
In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token-based
consistency model that generalizes a family of consistency models including
causal consistency, sequential consistency, and RedBlue consistency. An
application designer defines a set of tokens and specifies which pairs of
tokens conflict, and transactions acquire some subset of the tokens when they
execute. This allows the application designer to specify which transactions
conflict with each other.
%
% With no tokens, the model is equivalent to causal consistency. With all
% transactions acquiring a single self-conflicting token, the model is equivalent
% to sequential consistency~\cite{lamport1979make}. With some transactions
% acquiring a single self-conflicting token and some transactions acquiring no
% tokens, the model is equivalent to RedBlue consistency.
%
Gotsman et al.\ develop sufficient conditions to determine whether a given
token scheme is sufficient to guarantee that a global invariant is never
broken. This model of consistency is very similar to segmented
\invariantconfluence{}. The token-based approach allows users to specify certain
conflicts that are not possible with segmented \invariantconfluence{}. A
segmentation only allows transactions within a segment to acquire a single
self-conflicting lock. However, segmented \invariantconfluence{} also introduces
the notion of invariant segmentation, which cannot be emulated with the token
based approach. For example, it is difficult to emulate escrow transactions
with the token-based approach.

\textbf{CRDTs.}
CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
semilattices with inflationary update methods. Due to their algebraic
properties, CRDTs can be replicated with strong eventual consistency without
the need for any coordination. Our definition of distributed objects and our
\invariantconfluence{} system model are inspired directly by the corresponding
definitions and system models in~\cite{shapiro2011conflict}. Note that an
\invariantconfluent{} object does not necessarily have to be eventually
consistent, so it is natural (though not necessary) to use CRDTs as
\invariantconfluent{} distributed objects, achieving a combination of eventual
consistency and invariant preservation. Our criteria in
\thmref{LatticeProperty} also borrow ideas from CRDTs, exploiting the nice
algebraic properties of semilattices.

\textbf{CALM Theorem.}
Bloom~\cite{alvaro2010boom, alvaro2011consistency, conway2012logic} and its
formalism, Dedalus~\cite{alvaro2011dedalus, alvaro2013declarative}, are
declarative Datalog-based programming languages that are designed to program
distributed systems. The accompanying CALM
theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
a program can be written in the monotone fragment of these languages, then
there exists a coordination-free implementation of the program.
% Moreover, monotone programs are invariant to message reordering, whereas
% non-monotone programs may experience anomalies if messages are reordered in the
% network.
The CALM theorem provides guarantees about the outputs of a program, but it
does not guarantee that invariants are maintained throughout the duration of an
execution. Moreover, Bloom and Dedalus are general purpose programming
languages that can be used to implement a variety of distributed systems (e.g.
a sharded database) which are outside of the scope of \invariantconfluence{}.

@@

% \textbf{Proof Assistants.}
% Our interactive decision procedure bears resemblance to interactive proof
% assistants like Coq~\cite{coq2017}, Isabelle~\cite{nipkow2002isabelle},
% Agda~\cite{norell2008dependently}, and Idris~\cite{brady2013idris}. Proof
% assistants guide users through the process of interactively provided a proof
% that witnesses the validity of a formula. Our interactive decision procedure
% guides users through the process of interactively generating a set of reachable
% and unreachable points that witness the \invariantconfluence{} of an object. An
% interesting avenue for future work is to integrate an proof assistant into our
% interactive decision procedure. For example, a proof assistant could verify
% that a state labelled unreachable by a user is actually unreachable.

>>>>

\newcommand{\sieve}{\textsc{SIEVE}}
\textbf{RedBlue Consistency and \sieve.}
RedBlue consistency is a consistency model that sits between causal consistency
and linearizability~\cite{li2012making}.
%
With RedBlue consistency, every operation is manually labelled as either red or
blue. All operations are executed with causal consistency, but with the added
restrictions that red operations are executed in a single total order embedded
within the causal ordering.
%
Invariant safety is a sufficient (but not necessary) condition for RedBlue
consistent objects to be \invariantconfluent{} and is an analog of
\invariantclosure{} for an operation-based system model (as opposed to the
state based system model that we use).  In~\cite{li2014automating}, Li et al.\
develop sophisticated techniques for deciding invariant safety that involve
calculating weakest preconditions. These techniques are complementary to our
work and can be used to improve the \invariantclosure{} subroutine used by our
decision procedures.

\textbf{The Homeostasis Protocol.}
The homeostasis protocol~\cite{roy2015homeostasis} uses program analysis to
avoid unnecessary coordination between servers in a \emph{sharded} database
(whereas \invariantconfluence{} targets \emph{replicated} databases). The
protocol generates global invariants, called global treaties, and servers
operate without coordination so long as they preserve the treaty. When the
treaty is violated, the nodes perform a round of coordination and establish a
new treaty. The protocol guarantees that transactions are executed with
observational equivalence with respect to some serial execution of the
transactions. This means that intermediate states may be inconsistent, but
externally observable side effects and the final database state are consistent.
The observational equivalence guaranteed by the homeostasis is stronger than
the guarantees of \invariantconfluence{}. As a result, there are invariants and
workloads that the homeostasis protocol would execute with more coordination
than a segmented \invariantconfluent{} execution. Still, the homeostasis
protocol's mechanism of establishing invariants and operating without
coordination so long as the invariants are maintained is very similar to our
design of segmented \invariantconfluence{}.

@@

\textbf{Token Based Invariant Confluence.}
In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token based
consistency model that generalizes a family of consistency models including
causal consistency, sequential consistency, and RedBlue consistency. An
application designer defines a set of tokens and specifies which pairs of
tokens conflict, and transactions acquire some subset of the tokens when they
execute. This allows the application designer to specify which transactions
conflict with one another.  Gotsman et al.\ develop sufficient conditions to
determine whether a given token scheme is sufficient to guarantee that a global
invariant is never broken. This model of consistency is very similar to
segmented \invariantconfluence{}. The token based approach allows users to
specify certain conflicts that are not possible with segmented
\invariantconfluence{}. A segmentation only allows transactions within a
segment to acquire a single self-conflicting lock. However, segmented
\invariantconfluence{} also introduces the notion of invariant segmentation,
which cannot be emulated with the token based approach. For example, it is
difficult to emulate escrow transactions with the token based approach.

\textbf{CRDTs.}
CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
semilattices with inflationary update methods. Due to their algebraic
properties, CRDTs can be replicated with strong eventual consistency without
the need for any coordination. Our definition of distributed objects and our
\invariantconfluence{} system model are inspired directly by the corresponding
definitions and system models in~\cite{shapiro2011conflict}. Also, an
\invariantconfluent{} object does not necessarily have to be eventually
consistent, so it is natural (though not necessary) to use CRDTs as
\invariantconfluent{} distributed objects, achieving a combination of eventual
consistency and invariant preservation. Our criteria in
\thmref{LatticeProperty} also borrow ideas from CRDTs, exploiting the nice
algebraic properties of semilattices.

\textbf{CALM Theorem.}
Bloom~\cite{alvaro2010boom, alvaro2011consistency, conway2012logic} and its
formalism, Dedalus~\cite{alvaro2011dedalus, alvaro2013declarative}, are
declarative Datalog-based programming languages that are designed to program
distributed systems. The accompanying CALM
theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
a program can be written in the monotone fragment of these languages, then
there exists a coordination-free implementation of the program.  The CALM
theorem provides guarantees about the outputs of a program, but it does not
guarantee that invariants are maintained throughout the duration of an
execution. Moreover, Bloom and Dedalus are general purpose programming
languages that can be used to implement a variety of distributed systems (e.g.
a sharded database) which are outside of the scope of \invariantconfluence{}.

@@
